---
title: "namematcher"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{namematcher}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In data cleaning, you may need to compare names from different data sources to determine whether they belong to the same person or entity. For example:

- Establishing a person-level identifier across two data sources where names are given in different, unpredictable formats
- Matching people to their emails, usernames, or other name variants

The namematcher package provides tools for making these comparisons. This vignette introduces namematcher's first function, `js_divergence`, and demonstrates how it can be used in a basic workflow to infer name matches.

## Evaluate name similarity with `js_divergence`

The `js_divergence` function quantifies the token overlap between two names. In other words, to what extent are two names made up of different sub-components or parts?

It achieves this comparison by: 
1. Splitting the names into n-grams (slices of two letters, by default)
2. Calculating the probability that each name includes each shared n-gram
3. Calculating the Jensen-Shannon divergence on the mixture distribution of those probabilities

`js_divergence` expresses the difference between two names as a double between 0 and 1. The higher the value, the more dissimilar the names are from one another, and therefore the less likely they are to refer to the same individual.

### Basic use

```{r setup}
library(namematcher)

js_divergence("jon cooper", "jonathan cooper")
js_divergence("jon cooper", "joe collins")
```
