---
title: "namematcher"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{namematcher}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In data cleaning, you may need to compare names from different data sources to determine whether they belong to the same person or entity. For example:

- Establishing a person-level identifier across two data sources where names are given in different, unpredictable formats
- Matching people to their emails, usernames, or other name variants

The namematcher package provides tools for making these comparisons. This vignette introduces namematcher's first function, `js_divergence`, and demonstrates how it can be used in a basic workflow to infer name matches.

## Evaluate name similarity with `js_divergence`

The `js_divergence` function quantifies the token overlap between pairs of names using the Jensen-Shannon Divergence. In other words, to what extent do two names have overlapping sub-components?

It achieves this comparison by: 
1. Splitting the names into n-grams (slices of two letters, by default)
2. Getting n-grams shared by both names
3. Calculating the probability that each name includes each shared n-gram
4. Calculating the Jensen-Shannon Divergence on the mixture distribution of those probabilities

`js_divergence` expresses the difference between two names as a double between 0 and 1. The higher the value, the more dissimilar the names are from one another, and therefore the less likely they are to refer to the same individual.

### Basic use

`js_divergence` will return results character vectors of any length.

```{r basic-use}
library(namematcher)

# Evalute scalars. Returns 0: no divergence
js_divergence("Jon Smith", "Jon Smith")

# Calculates divergence pairwise across vectors
js_divergence(
  c("Jon Smith", "Elizabeth Howell"),
  c("Jon B Smith", "Leanne Holmes")
)
```

Because it handles vectors, it can be used in tidyverse syntax with ease. 

Below are some real names taken from the (DBPL Dataset 2)[https://hpi.de/naumann/projects/repeatability/datasets/dblp-dataset.html], with the column `sameentity` recording whether the names in question truly are the same individual. Using mutate, we can calculate the js_divergence between the `name_1` and `name_2` columns in a pipe sequence. This shows low JSDs for pairs who truly are the same individual, and high JSDs for pairs who are not.

```{r tidyverse-example}
library(dplyr)

sample_names <- tibble(
  name_1 = c("Said Hassan Ahmed", "Marcelino Bicho Dos Santos", "Stephen Fung", "Patrice Caire", "Detlef Sass"), 
  name_2 = c("Jagdish Chandra Patra", "Marcelino B. Santos", "Andrew E. Smith", "Kohtaro Ohba", "Detlef Saß"),
  sameentity = c(FALSE, TRUE, FALSE, FALSE, TRUE)
) |>
  mutate(js_divergence = js_divergence(name_1, name_2))

sample_names
```